---
title: "rMyCoPortal - Introduction"
author: "Franz-Sebastian Krah"
output: pdf_document
---

## Requirements
Under the new MacOS Mojave there have been problems with the sf and raster package. This vignette works fine under Sierra and Windows10. However, the vignette *application example* might create errors.

R version 3.5.1 (2018-07-02)
Platform: x86_64-apple-darwin15.6.0 (64-bit)
Running under: macOS High Sierra 10.13.6


# The MyCoPortal

The Mycology Collections data Portal (MyCoPortal) is a database of fungal diversity with records mainly from North America. For more - and detailled - information, please visit http://mycoportal.org/portal/index.php.

The rMyCoPortal R package is an interface to the content stored on the MyCoPoral website. It allows to download records from the database readily in R for further analysis. It further provides some basic plotting functions. Below I will show the basic usability and some further possibilites of using the data.

# Install rMyCoPortal
```{r setup, include=TRUE, eval=FALSE}

install.packages("devtools")
devtools::install_github("FranzKrah/rMyCoPortal")

```

## Docker

Before we start using rMyCoPortal, we need to install docker (https://docs.docker.com/install/). Docker performs  virtualization, also known as "containerization". rMyCoPortal interally uses the R package RSelenium to create a Selenium Server from which the MyCoPortal website is addressed. 
Docker needs to run before using the rMyCoPortal.

# Download records for *Amanita muscaria*, the fly agaric


```{r example1, include=TRUE, eval=TRUE, echo=TRUE}
## Load library
library("rMyCoPortal")

## Download records

am.rec <- mycoportal(taxon = "Amanita muscaria")
am.rec

head(am.rec@records)
```

## Visualization
We can now use several plotting methods to visualize the data.

```{r plots, include=TRUE, eval=TRUE, echo=TRUE}

x <- am.rec

## plot_recordstreemap can be used to visualize relative importance of aspects of the data
plot_recordstreemap(x = x, groupvar = "country", log = FALSE) # e.g., the country distribution

## plot_distmap can be used to plot interactive and static distribution maps
p1 <- plot_distmap(x = x, mapdatabase = "world", interactive = FALSE, plot = FALSE) # the default is interactive

# same for states
p2 <- plot_distmap(x = x, mapdatabase = "state", interactive = FALSE, plot = FALSE)

cowplot::plot_grid(p1, p2, ncol = 1, align = T)

## plot_datamap can be used to plot heatmaps for either records or species richness (index = "rich")
p3 <- plot_datamap(x = x, mapdatabase = "world", index = "rec", plot = FALSE)

## the same but cropped to Europe
p4 <- plot_datamap(x = x, mapdatabase = "state", index = "rec", plot = FALSE)

cowplot::plot_grid(p3, p4, ncol = 1, align = TRUE)

## And we can look up details for specific specimens
library(magick)
det <- details(x@records$Symbiota.ID[1])
length(det$urls)
par(mfrow = c(1,2), mar = c(0,0,0,0))
plot(image_read(det$urls[1]))
plot(image_read(det$urls[3]))

```
